---
title: "Homework 1"
author: "Mariel Gandarillas Calderon"
date: "2024-10-08"
output: html_document
---

```         
```

## Homework 1

## 1. Hankel Matrix Construction

1.1. Write a function `hankel_matrix` that computes a Hankel matrix of `dim_mat` rows and columns from a sequence `h_seq`. Use a `for` loop to generate the matrix.

```{r}
hankel_matrix <- function(dim_mat, h_seq) {
  required_length <- 2 * dim_mat - 1
  if (length(h_seq) < required_length) {
    stop("The length of h_seq must be at least 2 * dim_mat - 1.")
  }
  H <- matrix(0, nrow = dim_mat, ncol = dim_mat)
  for (i in 1:dim_mat) {
    for (j in 1:dim_mat) {
      H[i, j] <- h_seq[i + j - 1]
    }
  }
  
  return(H)
}
```

1.2 Modify the `hankel_matrix` function with input validation and informative messages. What is the minimum dimension N of the sequence h? Test cases:

```{r}
hankel_matrix <- function(dim_mat, h_seq) {
  if (!is.numeric(dim_mat) || length(dim_mat) != 1 || dim_mat <= 0 || dim_mat != as.integer(dim_mat)) {
    stop("Error: 'dim_mat' must be a positive integer(not decimals).")
  }
 
  if (!is.numeric(h_seq) || !is.vector(h_seq)) {
    stop("Error: 'h_seq' must be a numeric vector.")
  }
  
  required_length <- 2 * dim_mat - 1
  if (length(h_seq) < required_length) {
    stop("Error: 'h_seq' must have at least ", required_length, " elements to form a Hankel matrix of dimension ", dim_mat, "x", dim_mat, ".")
  }
  
  # Initialize an empty matrix
  H <- matrix(0, nrow = dim_mat, ncol = dim_mat)

  for (i in 1:dim_mat) {
    for (j in 1:dim_mat) {
      H[i, j] <- h_seq[i + j - 1]
    }
  }
  
  return(H)
}
```

3.  Alternative implementation without `for` loops: Write a function `hankel_matrix_fast` that returns the same Hankel matrix as `hankel_matrix` but avoids using a `for` loop. Notice that for a n×n Hankel matrix and a sequence h, we have the following: for the first row (or column) we have Hi,1=hi,i=1,…,n, for the second row (or column) we have Hi,2=hi,i=2,…,n+1, …, for the nth row (or column) we have Hi,n=hi,i=n,…,2n−1. Based on these observations, propose an alternative implementation using `rep` and `seq` functions, but without using `for` loops. Verify that the output of `hankel_matrix_fast` is the same as `hankel_matrix` by using `all.equal`.

    ```{r}
    hankel_matrix_fast <- function(dim_mat, h_seq) {
      # Input validation dim mat
      if (!is.numeric(dim_mat) || length(dim_mat) != 1 || dim_mat <= 0 || dim_mat != as.integer(dim_mat)) {
        stop("Error: 'dim_mat' must be a positive integer.")
      }
      
      # Input validation h_seq
      if (!is.numeric(h_seq) || !is.vector(h_seq)) {
        stop("Error: 'h_seq' must be a numeric vector.")
      }

      required_length <- 2 * dim_mat - 1
      if (length(h_seq) < required_length) {
        stop(paste("Error: 'h_seq' must have at least", required_length, "elements to form a Hankel matrix of dimension", dim_mat, "x", dim_mat, "."))
      }
      
      # Generate row and column indices with 'rep' 
      row_indices <- rep(1:dim_mat, times = dim_mat)
      col_indices <- rep(1:dim_mat, each = dim_mat)
      
      indices <- row_indices + col_indices - 1
      
      # Create the Hankel matrix 
      H <- matrix(h_seq[indices], nrow = dim_mat, ncol = dim_mat)
      
      return(H)
    }
    ```

4.  Benchmarking: Benchmark your two implementations. Use `microbenchmark::microbenchmark({...},{...})` where `...` is to be replaced by your implementations. Comment the results. The vectorized version (`hankel_matrix_fast`) is expected to be faster. Make a graph in order to compare benchmark results with respect to growing dimensions `dim_mat`.

```{r}

```
